---
title: "Analysis of Model Validation Results"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(results = "markup", fig.align = "center",
                      fig.height= 8, fig.width= 8,message=FALSE,warning=FALSE)
```

# Introduction

This part of the analysis follows the predictive modeling (implemented using Python in jupyter notebooks). Based on the performance in the validation set, we selected Light Gradient Boosting (lgbm) as the best performing model after tuning its performance. Here our goal is to derive conclusions from our model to better understand its capacity and limitations.

# Loading validations results

Validation results were saved as a flat file where the validation data set was linked to model predictions as well as the true decision time and absolute error:

```{r, eval=FALSE}
validation_results <- read.csv("validation_results.csv", stringsAsFactors = F, header = T)
```

We also get our master data file:

```{r,eval= FALSE}
# Load previously save intermediary data file
current <- readRDS("current.rds")
```


We will append validation results with additional information from the master data:

```{r,eval=FALSE}
validation_results_extended <- merge(x = validation_results,y = current, by = c('KNUMBER','APPLICANT_PRIOR_CLEARANCE_TO_DATE','DEVICENAME_PRIOR_CLEARANCE_TO_DATE','DECISIONTIME'), all.x = TRUE, all.y = FALSE)
```


# Converting Medical Speciality codes to identifiers

At this point let's also convert speciality codes to more informative forms. These codes are found at:

https://www.fda.gov/medicaldevices/deviceregulationandguidance/overview/classifyyourdevice/ucm051668.htm#medicalspecialty

Downloaded as a flat file (MedSpeciality.csv) and used:

```{r}
medspeciality <- read.csv("MedSpeciality.csv", stringsAsFactors = F, header = T)
names(medspeciality)[1] <- "Speciality"
medspeciality$Speciality <- gsub(pattern = " |&|/|,","_",medspeciality$Speciality)
medspeciality$Speciality <- gsub(pattern = "__|___","_",medspeciality$Speciality)
```

```{r}
validation_results_extended$REVIEWADVISECOMM <- as.character(validation_results_extended$REVIEWADVISECOMM)
validation_results_extended$CLASSADVISECOMM <- as.character(validation_results_extended$CLASSADVISECOMM)
for(i in seq_along(medspeciality$Medical.Specialty.Code)){
    scode <- medspeciality$Medical.Specialty.Code[i]
    selector_REVIEWADVISECOMM <- which(validation_results_extended$REVIEWADVISECOMM == scode)   
    selector_CLASSADVISECOMM <- which(validation_results_extended$CLASSADVISECOMM == scode)
    validation_results_extended$REVIEWADVISECOMM[selector_REVIEWADVISECOMM] <- medspeciality$Speciality[i]
    validation_results_extended$CLASSADVISECOMM[selector_CLASSADVISECOMM] <- medspeciality$Speciality[i]
}
```


This extended results table serves as the main data set for our validation results analysis.

```{r,eval= FALSE}
saveRDS(validation_results_extended,"validation_results_extended.rds")
```



# Overall Model Performance


```{r}
validation_results_extended <- readRDS("validation_results_extended.rds")
```


```{r, fig.height= 5, fig.width= 5, fig.align= "center"}
mae = round(median(validation_results_extended$Absolute_Error),4)
library(ggplot2)
ggplot(data = validation_results_extended, aes(x = preds_lgbm, y = DECISIONTIME))+
    geom_point(alpha = 0.05, color = "navy", size = 1)+
    geom_abline(x = c(0,0), y = c(1000,1000), linetype = "dashed", size = 1, color = "magenta")+
    geom_text(x = 500, y = 900, label = paste0("MAE: ", mae), family = "arial", size = 6, color = "navy")+
    ylim(0,1000)+
    xlim(0,1000)+
    xlab(label = "Model predictions (calendar days)")+
    ylab(label = "Decision time (calendar days)")+
    ggtitle("Validated model performance")+
    theme(
        axis.text = element_text(size = 12, face = "bold", colour = "navy"),
        axis.title = element_text(size = 15, face = "bold", colour = "navy"),
        plot.title = element_text(size = 18, face = "bold", colour = "navy", hjust = 0.5),
        panel.background = element_blank(),
        panel.border = element_rect(color = "navy", fill = NA),
        panel.grid.major = element_line(size = 0.05, colour = "grey", linetype = "dotted")
    )

    
```

# Error distribution

```{r,fig.height= 5, fig.width= 5, fig.align= "center"}
library(ggplot2)
quant <- round(quantile(validation_results_extended$Absolute_Error),2)
ggplot(validation_results_extended,aes(x = Absolute_Error))+
    geom_histogram(bins =  1000, color= "navy", size = 0.05)+
    xlab(label = "Absolute Error (calendar days)")+
    ylab(label = "Count")+
    ggtitle("Distribution of Validation Errors")+
    geom_vline(xintercept = quant[3], size = 0.75, linetype = "dashed", color = "magenta")+
    geom_text(x = 180, y = 500, label = paste0("MAE: ", mae), family = "arial", size = 6, color = "navy")+
    xlim(-100,800)+
    ylim(0,500)+
    theme(
        axis.text = element_text(size = 12, face = "bold", colour = "navy"),
        axis.title = element_text(size = 15, face = "bold", colour = "navy"),
        plot.title = element_text(size = 18, face = "bold", colour = "navy", hjust = 0.5),
        panel.background = element_blank(),
        panel.border = element_rect(color = "navy", fill = NA)
        #panel.grid.major = element_line(size = 0.05, colour = "grey", linetype = "dotted")
    )
```
```{r,fig.height= 5, fig.width= 5, fig.align= "center"}
library(ggplot2)
quant <- round(quantile(validation_results_extended$Absolute_Error),2)
ggplot(validation_results_extended,aes(y = Absolute_Error, x= DECISIONTIME))+
    geom_point(alpha = 0.015, color = "navy")+
    ylab(label = "Absolute Error (calendar days)")+
    xlab(label = "Decision time (calendar days)")+
    geom_vline(xintercept = 150, color = "red", linetype = "dashed", size = 1.0)+
    geom_text(y = 600, x = 300, label = "150 Days",  size = 5, color = "red")+
    geom_hline(yintercept = quant[3], color = "magenta", linetype = "dashed", size = 1.2)+
    geom_text(y = 70, x = 750, label = paste0("MAE: ", quant[3]),  size = 6, color = "navy")+
    theme(
        axis.text = element_text(size = 12, face = "bold", colour = "navy"),
        axis.title = element_text(size = 15, face = "bold", colour = "navy"),
        plot.title = element_text(size = 18, face = "bold", colour = "navy", hjust = 0.5),
        panel.background = element_blank(),
        panel.border = element_rect(color = "navy", fill = NA),
        panel.grid.major = element_line(size = 0.05, colour = "grey", linetype = "dotted")
    )
```

Error distribution is generally uniform until 150 decision days. Submissions that are taking longer than 150 days are associated with higher prediction errors, suggesting that there are other variables that model could not explain.


# Analysis of model performance by medical specialities

```{r, fig.width=10}
library(dplyr)
rev_spec_sum <- validation_results_extended %>% group_by(REVIEWADVISECOMM) %>% summarise(Median_prediction_error = median(Absolute_Error), Median_decision_time = median(DECISIONTIME),Stdev_decision_time = sd(DECISIONTIME), Number_of_submissions = n())
rev_spec_sum$REVIEWADVISECOMM <- gsub("__|_"," ",rev_spec_sum$REVIEWADVISECOMM)
library(ggplot2)
fac <- with(rev_spec_sum, reorder(REVIEWADVISECOMM, Median_prediction_error, order = TRUE))
rev_spec_sum$REVIEWADVISECOMM <- factor(rev_spec_sum$REVIEWADVISECOMM, levels = levels(fac))
library(reshape2)
test <- melt(rev_spec_sum,id.vars = "REVIEWADVISECOMM")
fac_levels <- levels(test$variable)
test$variable <- factor(gsub("__|_"," ",test$variable), levels = gsub("__|_"," ",fac_levels))
ggplot(test, aes(x= REVIEWADVISECOMM, y = value))+
    geom_bar(stat = "identity", fill = "navy")+
    facet_grid(. ~ variable,scales = "free_x")+
    xlab("Review Advisery Medical Speciality\n")+
    ylab("")+
    theme(
        axis.text = element_text(size = 10, face = "bold", colour = "navy"),
        axis.title = element_text(size = 15, face = "bold", colour = "navy"),
        plot.title = element_text(size = 18, face = "bold", colour = "navy", hjust = 0.5),
        panel.background = element_blank(),
        panel.border = element_rect(color = "navy", fill = NA),
        panel.grid.major.x = element_line(size = 0.05, colour = "navy", linetype = "dotted"),
        strip.background = element_rect(fill = "purple4"),
        strip.text = element_text(color = "white", size = 12)
    )+
    coord_flip()
  
```

# Analysis of model performance by product codes

```{r, fig.width=10}
library(dplyr)
prod_code_sum <- validation_results_extended %>% group_by(PRODUCTCODE) %>% summarise(Median_prediction_error = median(Absolute_Error), Median_decision_time = median(DECISIONTIME),Stdev_decision_time = sd(DECISIONTIME), Number_of_submissions = n())

# Filter at least 10 submissions or more
prod_code_sum <- filter(prod_code_sum, Number_of_submissions >= 10)

# 20 Product codes with the least median prediction error
prod_code_sum <- prod_code_sum[order(prod_code_sum$Median_prediction_error),]
prod_code_sum20 <- prod_code_sum[1:20,]

# Attach product code descriptions
# Dowloaded from: https://www.fda.gov/medicaldevices/deviceregulationandguidance/overview/classifyyourdevice/ucm051668.htm

pcodes <- read.delim("foiclass.txt", sep = "|", header = T, stringsAsFactors = F)
select_pcodes <-  pcodes$PRODUCTCODE %in% prod_code_sum20$PRODUCTCODE 
pcodes <- pcodes[select_pcodes,c("PRODUCTCODE","DEVICENAME")]
#pcodes$DEVICENAME <- paste0("(",pcodes$DEVICENAME,")")

prod_code_sum20 <- merge(prod_code_sum20,pcodes,by = "PRODUCTCODE", sort = FALSE)
prod_code_sum20$PRODUCTCODE <- paste(prod_code_sum20$PRODUCTCODE,prod_code_sum20$DEVICENAME,sep = " \n")
prod_code_sum20 <- dplyr::select(prod_code_sum20,-DEVICENAME)

prod_code_sum20$PRODUCTCODE <- sapply(prod_code_sum20$PRODUCTCODE,function(x){return(substr(x,1,60))}) 

library(ggplot2)
fac <- with(prod_code_sum20, reorder(PRODUCTCODE, Median_prediction_error, order = TRUE))
prod_code_sum20$PRODUCTCODE <- factor(prod_code_sum20$PRODUCTCODE, levels = levels(fac))
library(reshape2)
test <- melt(prod_code_sum20,id.vars = "PRODUCTCODE")
fac_levels <- levels(test$variable)
test$variable <- factor(gsub("__|_"," ",test$variable), levels = gsub("__|_"," ",fac_levels))

ggplot(test, aes(x= PRODUCTCODE, y = value))+
    geom_bar(stat = "identity", fill = "navy")+
    facet_grid(. ~ variable,scales = "free_x")+
    xlab("Product Code\n")+
    ylab("")+
    theme(
        axis.text.x = element_text(size = 11, face = "bold", colour = "navy"),
        axis.text.y = element_text(size = 8, face = "bold", colour = "navy", hjust = 0.5),
        axis.title = element_text(size = 15, face = "bold", colour = "navy"),
        plot.title = element_text(size = 18, face = "bold", colour = "navy", hjust = 0.5),
        panel.background = element_blank(),
        panel.border = element_rect(color = "navy", fill = NA),
        panel.grid.major.x = element_line(size = 0.05, colour = "navy", linetype = "dotted"),
        strip.background = element_rect(fill = "purple4"),
        strip.text = element_text(color = "white", size = 10)
    )+
    coord_flip()
  
```

# Plotting data for a given product code

It would be useful to add open names of product codes from the product code table into our extended validation results data:


```{r,eval=FALSE}
pcodes <- read.delim("foiclass.txt", sep = "|", header = T, stringsAsFactors = F)
pcodes <- pcodes[c("PRODUCTCODE","DEVICENAME")]
names(pcodes)[2] <- "PRODUCTCODENAME"
validation_results_extended <- merge(validation_results_extended,pcodes, by = "PRODUCTCODE", sort = FALSE, all.x = T, all.y = F)
saveRDS(validation_results_extended,"validation_results_extended.rds")
```

Use the updated table with product code name column added: 

```{r}
validation_results_extended <- readRDS("validation_results_extended.rds")
```




```{r}
# Function to scatterplot prediction results along with real decision time from a single product code. 
pcode <- "GZB"
results <- "validation_results_extended"

plot_productcode <- function(results,pcode){
    require(dplyr)
    require(ggplot2)
    subdata <- dplyr::filter(results, PRODUCTCODE == pcode)
    mae = round(median(subdata$Absolute_Error),2)
    edge = max(c(max(subdata$DECISIONTIME),max(subdata$preds_lgbm)))
    ggplot(data = subdata, aes(x = preds_lgbm, y = DECISIONTIME))+
        geom_point(color = "navy", size = 4)+
        geom_abline(x = c(0,0), y = c(1000,1000), linetype = "dashed", size = 1, color = "magenta")+
        geom_text(x = 10, y = edge, label = paste0("MAE: ", mae), size = 5, color = "navy")+
        ylim(0, edge + 10)+
        xlim(0, edge + 10)+
        xlab(label = "Model predictions (calendar days)")+
        ylab(label = "Real Decision time (calendar days)\n")+
        labs(title = "Validated model performance", 
             subtitle = paste0("Product code: |",pcode,"| ",subdata$PRODUCTCODENAME[1]))+
        theme(
            axis.text = element_text(size = 12, face = "bold", colour = "navy"),
            axis.title = element_text(size = 15, face = "bold", colour = "navy"),
            plot.title = element_text(size = 18, face = "bold", colour = "navy", hjust = 0.5),
            plot.subtitle = element_text(size = 12, face = "bold", colour = "navy", hjust = 0.5),
            panel.background = element_rect(fill = "slategray1"),
            panel.border = element_rect(color = "navy", fill = NA),
            panel.grid.major = element_line(size = 0.05, colour = "grey", linetype = "dotted"),
            panel.grid.minor = element_blank()
        )
    
}

plot_productcode(validation_results_extended,pcode)

```

```{r}
# Function to prepare timeseries plot of prediction results along with real decision time from a single product code. 
pcode <- "GZB"
results <- "validation_results_extended"

plot_timeseries_productcode <- function(results,pcode){
    require(dplyr)
    require(ggplot2)
    require(reshape2)
    subdata <- dplyr::filter(results, PRODUCTCODE == pcode)
    subdata$DATERECEIVED <- as.POSIXct.Date(subdata$DATERECEIVED)
    mae = round(median(subdata$Absolute_Error),2)
    edge = max(c(max(subdata$DECISIONTIME),max(subdata$preds_lgbm)))
    
    ggplot(data = subdata, aes(x = DATERECEIVED))+
        geom_segment(aes(x = DATERECEIVED, y = DECISIONTIME), 
                     yend = subdata$preds_lgbm, xend = subdata$DATERECEIVED,
                     linetype = "dotted", size = 1.0, color = "red",show.legend = T)+
        geom_point(color = "magenta", size = 4, aes(x = DATERECEIVED, y = DECISIONTIME) , alpha = 0.85, show.legend = T)+
        geom_point(color = "navy", size = 4, aes(x = DATERECEIVED, y = preds_lgbm), alpha = 0.85, show.legend = T)+
        

        xlab(label = "Submission Received Date")+
        ylab(label = "Decision time or Predictions \n(calendar days)")+
        labs(title = "Decision time and predictions across time", 
             subtitle = paste0("Product code: |",pcode,"| ",subdata$PRODUCTCODENAME[1]))+
        theme(
            axis.text = element_text(size = 12, face = "bold", colour = "navy"),
            axis.title = element_text(size = 15, face = "bold", colour = "navy"),
            plot.title = element_text(size = 18, face = "bold", colour = "navy", hjust = 0.5),
            plot.subtitle = element_text(size = 12, face = "bold", colour = "navy", hjust = 0.5),
            panel.background = element_rect(fill = "slategray1"),
            panel.border = element_rect(color = "navy", fill = NA),
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank()
           # panel.grid.major = element_line(size = 0.05, colour = "grey", linetype = "dotted")
        )
    
}

plot_timeseries_productcode(validation_results_extended,pcode)

```